What Happened to The Birth Rate?

### 2022-02-28 outputs/paper/paper.Rmd
Changed minor details, typos, wording

---
title: "What Happened to The Birth Rate?"
subtitle: "The Great Recession and the LGBT Movement"
author: 
  - Lantao Cui
  - Huakun Shen
  - Sijian Wu
thanks: "Replication avaliable at https://doi.org/10.48152/ssrp-mbxx-jm47. Code and data avaliable at https://github.com/cuilantao/Investigation-of-falling-US-Birth-Rates-since-the-great-recession"         
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "This paper is based on Kearney's paper \"The Puzzle of Falling US Birth Rates since the Great Recession\". ^[Original paper available at: https://www.aeaweb.org/articles?id=10.1257/app.2.2.95] [@cite_original] During the Great Recession, the US birth rate varied between a tight range of roughly 65 to 70 birth per 1000 women between the age of 15 to 44. However, after the Great Recession, the birth rate started to decline significantly, it dropped by nearly 20 percent in 13 years. In this paper, we will expand our research on multiple countries and find out whether the level of impact from recession would be correlated with the changes in the birth rate for different countries. We will also treat the rise in the LGBT community as a potential cause for the decline in birth rate and examine its effect"
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
header-includes:
  - \usepackage{subfig}
  - \usepackage{float}
  - \floatplacement{figure}{H}
---

@@

In the United States, unemployment rates, mortgage foreclosures, and poverty rates rose while housing values fell, but the extent of these changes varied widely across local areas. Another impact from the depressed economic condition is delayed childbearing, and others to hasten it. The fertility rate declined at the national level, dropping from a recent high in 2007 of 69.5 births per 1,000 women aged 15 to 44, to 63.2 for 2012.1 There was, however, great variation by state, age, and ethnicity, with younger and Hispanic women showing disproportionate decreases[@fertility]. The significant consequence on fertility rate would make us wonder how severe is America being affected by the great recession. However,  by comparing currency devaluations, equity market declines, and rising sovereign bond spreads, the U.S. is the third less affected country[@rank].

@@

Another thing that might have affected the fertility rate is the rise of the LGBT community, the LGBT activity started in the 1920s and had huge progress after the 2000s. A lot of laws that prohibited homosexual relationships were lifted and homosexual couples can get married legally and adopt kids, there is also propaganda that encourages homosexuality.

@@

## Data Source
Part of our data are from the NCHS(National Center for Health Statistics)[@nvss], in particular, we acquired the US birth rate data for the past 40 years, Figure \@ref(fig:fig-past-40-years) shows the US birth rate from 1980 - 2020, as we can see, the birth rate declined significantly after 2007, which is the time of the Great Recession. We also generated data from U.S. Birth Rate 1950-2022[@birthrate]. We selected the top 3 countries that are affected most and least by the Great Recession according to rankings determined by comparing currency devaluations, equity market declines, and rising sovereign bond spreads. These measures tend to track developments in the real economy during times of economic crisis when financial strain handicaps consumption, investment, and in many cases government spending, which limits GDP and employment growth[@rank]. TODO: ADD DATA SOURCE AND BREIFLY TALK ABOUT IT

@@

@@

Figure \@ref(fig:least-affected) shows the changes of births per 1000 people of China, Japan, the U.S, and South Africa since 1950 with the prediction for the future. The births rates of China and South Africa were obviously higher than the U.S. and Japan. However, around the year of recession, the birth rate of China and Japan dropped even harder than the U.S. on the absolute level that is lower than 10 births per 1000 people.

@@

Figure \@ref(fig:most-affected) shows the changes of births per 1000 people of Ukraine, Argentina, Hungary, Poland since 1950 with prediction for the future. All these most affected countries have lower initial births rate than those least affected countries. However, the levels of births per 1000 people were similar to those least affected countries around recession except Argentina. Ukraine, Hungary, and Poland also showed a potential increase after recession but continued to decline afterwards.

@@

Figure \@ref(fig:fig-parity) is a replication of a plot showing trends in births by "parity" (referring to the number of children for a given woman) from [@cite_original]. First and second births declined a lot since 2007, but the trend lines for the third and higher order births are comparatively flat.
The women giving their 3rd and 4th births are most likely older cohorts comparing to the cohorts giving their 1st and 2nd births. Again, this is another evidence showing the impact brought by cohort effects.

>>>>

---
title: "What Happened to The Birth Rate?"
subtitle: "The Great Recession and the LGBT Movement"
author: 
  - Lantao Cui
  - Huakun Shen
  - Sijian Wu
thanks: "Replication avaliable at https://doi.org/10.48152/ssrp-mbxx-jm47. Code and data avaliable at https://github.com/cuilantao/Investigation-of-falling-US-Birth-Rates-since-the-great-recession"         
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "This paper is based on Kearney's paper \"The Puzzle of Falling US Birth Rates since the Great Recession\". ^[Original paper available at: https://www.aeaweb.org/articles?id=10.1257/app.2.2.95] [@cite_original] During the Great Recession, the US birth rate varied between a tight range of roughly 65 to 70 births per 1000 women between the age of 15 to 44. However, after the Great Recession, the birth rate started to decline significantly, it dropped by nearly 20 percent in 13 years. In this paper, we will expand our research on multiple countries and find out whether the level of impact from recession would be correlated with the changes in the birth rate for different countries. We will also treat the rise in the LGBT community as a potential cause for the decline in birth rate and examine its effect."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
header-includes:
  - \usepackage{subfig}
  - \usepackage{float}
  - \floatplacement{figure}{H}
---

@@

In the United States, unemployment rates, mortgage foreclosures, and poverty rates rose while housing values fell, but the extent of these changes varied widely across local areas. Another impact from the depressed economic condition is delayed childbearing, and others to hasten it. The fertility rate declined at the national level, dropping from a recent high in 2007 of 69.5 births per 1,000 women aged 15 to 44, to 63.2 for 2012. There was, however, great variation by state, age, and ethnicity, with younger and Hispanic women showing disproportionate decreases[@fertility]. The significant consequence on fertility rate would make us wonder how severe is America being affected by the great recession. However,  by comparing currency devaluations, equity market declines, and rising sovereign bond spreads, the U.S. is the third less affected country[@rank].

@@

Another factor other than policy or economic factors that might have affected the fertility rate is cohort effect, more specifically, in this paper we focus on the rise of the LGBT community. The LGBT activity started in the 1920s and had a huge progress after the 2000s[@same_sex_marriage_states_2022]. A lot of laws that prohibited homosexual relationships were lifted and homosexual couples can get married legally and adopt kids, there is also propaganda that encourages homosexuality.

@@

## Data Source
Part of our data are from the NCHS(National Center for Health Statistics)[@nvss], in particular, we acquired the US birth rate data for the past 40 years, Figure \@ref(fig:fig-past-40-years) shows the US birth rate from 1980 - 2020, as we can see, the birth rate declined significantly after 2007, which is the time of the Great Recession. We also generated data from U.S. Birth Rate 1950-2022[@birthrate]. We selected the top 3 countries that are affected most and least by the Great Recession according to rankings determined by comparing currency devaluations, equity market declines, and rising sovereign bond spreads. These measures tend to track developments in the real economy during times of economic crisis when financial strain handicaps consumption, investment, and in many cases government spending, which limits GDP and employment growth[@rank].

The data used to analyze the impact to birth rate brought by LGBT population growth comes from the School of Law, UCLA[@will_institute]. The analysis was based on the data collected for the Gallup Daily Tracking survey.

The data used to analyze LGBT population trends comes from GALLUP News [@us_lgbt_trend], which is the source of data of the LGBT Demographics analysis[@will_institute]. The data is based on telephone interviews conducted in 2020 with a random sample of 15,349 adults (18+), living in the 50 states and the District of Columbia. The entire sample of national adults has a margin of sampling error is $\pm1\%$ at the $95\%$ confidence level. The results based on the total sample of 692 LGBT adults has a sample error of $\pm5\%$ with a $95\%$ confidence level.

@@

The LGBT demographics data spans 2012 to 2017. The sample is approximately 350,000 U.S. adults (18+) who resides in the 50 states and district of Columbia. The data describes LGBT Data & Demographics such as LGBT proportion of population by states, characteristics of LGBT people (race, gender distribution, etc.), socioeconomic indicators (income, education level, etc.) and state comparison.

The GALLUP News data used to analyze LGBT population trend shows the LGBT population proportion and distribution trend from 2012 to 2020 by year and generation, as well as the distribution of sexual orientation of different gender identities.

@@

Figure \@ref(fig:least-affected) shows the changes of births per 1000 people of China, Japan, the U.S, and South Africa since 1950 with the prediction for the future. The birth rates of China and South Africa were obviously higher than that of the U.S. and Japan. However, around the year of recession, the birth rate of China and Japan dropped even harder than the U.S. on the absolute level that is lower than 10 births per 1000 people.

@@

Figure \@ref(fig:most-affected) shows the changes of births per 1000 people of Ukraine, Argentina, Hungary, Poland since 1950 with prediction for the future. All these most affected countries have lower initial birth rates than that of the those least affected countries. However, the levels of births per 1000 people were similar to those least affected countries around recession except Argentina. Ukraine, Hungary, and Poland also showed a potential increase after recession but continued to decline afterwards.

@@

Figure \@ref(fig:fig-parity) is a reproduction of a plot showing trends in births by "parity" (referring to the number of children for a given woman) from [@cite_original]. The first and second births declined a lot since 2007, but the trend lines for the third and higher order births are comparatively flat.
The women giving their 3rd and 4th births are most likely older cohorts comparing to the cohorts giving their 1st and 2nd births. Again, this is another evidence showing the impact brought by cohort effects.
